NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-06-22 paper/application.tex
update paper

Interestingly, all configurations except one are concentrated around the full \emph{neutral} prediction. Referring back to the pipeline visualization, we observe that the only configuration that failed to produce the correct label is the one for which we allow only updating of the attention stage of the model.
%
%In other words, the updating of the attention stage seems to have significantly less impact on the prediction result compared to the classifier or encoder stage of the model.
%Similar observations can be obtained for most examples we have experimented with.
%\shusen{
%The domain experts comment that this observation is interesting and worth further investigation from the perspective of NLP research.
The domain experts find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From this analysis, we can extrapolate that adjusting the word alignment affects only the composed semantics but has no effect on word semantics, which often plays a significant role in a good model (as supported by recent works in word embeddings, e.g., ELMo~\cite{salant2017contextualized}). Without proper semantics (in both encoder and classifier), only adjusting alignment is not sufficient to yield good model. 
%To interpret the result, we need to carry out a case by case analysis, as the separation of the stages depends on the individual implementation.
%In our case, the result indicates that the tunable parameters of the attention stage (the transformation before the alignment operation, see details in Appendix A) has less impact on the final prediction compared to the other two stages.
%However, this result does not necessarily imply the limitation of attention mechanism as the core alignment computation does not have tunable parameters that can be modified by the proposed optimization.
%}
%The attention layer is less sensitive compared to encoding layer and classification layer.
%When examined other sentences pairs, such an observation remains.
%This gives domain expert some hint about the significance of the role of each pipeline stage play in this model.
% which based on the natural extension of the margin-infused relaxed algorithm (MIRA) to the neural network generalize the concept of maligin combined effort of the  proposed technique

>>>>

Interestingly, all configurations except one are concentrated around the full \emph{neutral} prediction. Referring back to the pipeline visualization, we observe that the only configuration that failed to produce the correct label is the one for which we allow only updating of the attention stage of the model.
%
%In other words, the updating of the attention stage seems to have significantly less impact on the prediction result compared to the classifier or encoder stage of the model.
%Similar observations can be obtained for most examples we have experimented with.
%\shusen{
%The domain experts comment that this observation is interesting and worth further investigation from the perspective of NLP research.
%The domain experts find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From this analysis, we can extrapolate that adjusting the word alignment affects only the composed semantics but has no effect on word semantics, which often plays a significant role in a good model (as supported by recent works in word embeddings, e.g., ELMo~\cite{salant2017contextualized}). Without proper semantics (in both encoder and classifier), only adjusting alignment is not sufficient to yield good model. 
%
The domain experts find this observation very interesting and suggest a preliminary interpretation. They believe the attention provides a way to compose word semantics. Individual word semantics are yielded from the encoder and input embedding, while the composed semantics participate in classification layer. From this analysis, we can extrapolate that the ultimate decision relies on the semantics (i.e., encodings and composed encodings in classifier). The encoder and classifier layer can swiftly adjust weights for word/composed semantics to correct prediction. But the attention layer, which only works on how individual words are aligned, affects the prediction less significantly. Recent works also support this observation in word embeddings, e.g., ELMo~\cite{salant2017contextualized}, where better word embeddings can substantially benefit a model. However, it by no means implies attention is not useful, as it serves as a way to compose word semantics.
%To interpret the result, we need to carry out a case by case analysis, as the separation of the stages depends on the individual implementation.
%In our case, the result indicates that the tunable parameters of the attention stage (the transformation before the alignment operation, see details in Appendix A) has less impact on the final prediction compared to the other two stages.
%However, this result does not necessarily imply the limitation of attention mechanism as the core alignment computation does not have tunable parameters that can be modified by the proposed optimization.
%}
%The attention layer is less sensitive compared to encoding layer and classification layer.
%When examined other sentences pairs, such an observation remains.
%This gives domain expert some hint about the significance of the role of each pipeline stage play in this model.
% which based on the natural extension of the margin-infused relaxed algorithm (MIRA) to the neural network generalize the concept of maligin combined effort of the  proposed technique
